A Domain Specific Language for SpectreGuard

### 2020-12-11 paper/main.tex
Intro; background; related work; future work

\section{Introduction}
SpectreGuard is provides an abstraction that allows programmers to prevent
speculative access of certain variables.~\cite{SpectreGuard} For
preserving program correctness, static analyses are desirable to prevent
programmers from accidentally successfully compiling code which copies a
``secret'' variable (a variable protected by SpectreGuard) to a ``public''
variable (a variable with no such protection). To this end, two interrelated static analysis techniques are designed and
implemented.

@@

\section{Background}
Cite a paper. % \cite{barroso2009datacenter}.

Cite multiple papers. % \cite{banga99resourcecontainers,barroso2009datacenter}

@@

@@

At the end of the algorithm, only leaves corresponding to public names are
retained.

\section{Related Work}
The way that the type-level tagging mechanism works is similar to several other
prior works. One such work is McBride's "Kleisli Arrows of Outrageous Fortune"~\cite{KleisliArrows}. In that paper, an invalid state is unable to be reached as the result of
a similar type-level tag as the \ttt{Sensitivity} tag in the present paper.
However, significantly more machinery is involved in the McBride paper, as
it is necessary to put such a tag on a monadic type. While a monadic type is
used in the present paper (in the form of the \ttt{Cmd} type), this type
does not use a type-level tag. The creation of such machinery is the
main purpose of the McBride paper.

\section{Future Work}
A further investigation into the tradeoffs and possible additional features of
the information flow analysis is likely to be useful. This could provide a more
accurate and succinct picture to programmers of where leaks are occurring, or
potentially occurring. It is possible that some of these other leak conditions
could be incorporated into the type system, by exploiting the dependent type
features enabled by singletons~\cite{SingletonsPaper} and, in the future, by
Dependent Haskell.~\cite{DepHaskSpec} Dependent types provide a powerful
technique for writing programs at a type-level, which will be executed at
compile-time by the type checker.~\cite{CertProg, DepHaskSpec} A more
lightweight technique, which might be able to be applied to allow for more
expressive policies to control of information flow within the type system, would
be the ``ghosts of departed proofs'' design technique.~\cite{DepartedProofs}

@@

\section{Conclusion}

>>>>

\section{Introduction}
\label{sec:Introduction}
SpectreGuard is provides an abstraction that allows programmers to prevent
speculative access of certain variables.~\cite{SpectreGuard} As Spectre-based
attacks remain an issue, even in modern computer systems, such mitigations and
preventative measures are crucial to maintaining the security of software.  For
preserving program correctness, static analyses are desirable to prevent
programmers from accidentally successfully compiling code which copies a
``secret'' variable (a variable protected by SpectreGuard) to a ``public''
variable (a variable with no such protection). To this end, two interrelated
static analysis techniques are designed and implemented.

@@

\section{Background}

Haskell is a purely functional programming language with strong static type system
based on System F and algebraic datatypes (ADTs).~\cite{HistoryOfHaskell} These features,
particularly the type system and the algebraic datatypes, make it a powerful tool
for the creation of EDSLs. Expressions, commands and programs (compound commands)
can be straightforwardly expressed using ADTs using a \textit{deep embedding}.~\cite{DeepAndShallow}

A deep embedding directly exposes its abstract syntax tree (AST) as a value.
This allows EDSL programs to be easily manipulated and examined, such as for
interpretation, code generation and static analysis (the latter two we make use
of in this paper).

@@

The one major aspect of the C code generator that is not described in this table
is how the loop variable in \ttt{for} loops can be tagged as secret (if appropriate).
This is accomplished by recovering the type-level sensitivity type information at
runtime. Such a recovery is made possible by the use of singletons.~\cite{Singletons}

@@

An alternate way to view the scope stack is as a stack of \textit{relations} on
on names. Each item of the stack is a relation such that two names are related
if and only if the left-hand name is a public variable declared in the corresponding
scope and the right-hand name is a secret variable in a conditional within the scope.

This allows us to look for leaks that happen ``through'' a control construct.

The operations on the scope stack data structure have the following type signatures:

\begin{lstlisting}
emptyScope :: Scope

scopeAddName :: Scope -> Name -> Scope
scopePush :: Scope -> [Name] -> Scope
scopePop :: Scope -> Scope
publicNonLocals :: Scope -> Name -> Maybe [Name]
\end{lstlisting}

These operations are used by the tool in its analysis of an EDSL
program as follows:

Whenever a declaration is encountered in EDSL code, the corresponding variable
name is added to the scope stack with \ttt{scopeAddName}. Whenever a control
structure is entered (such as a \ttt{for} loop or an \ttt{if} statement),
\ttt{scopePush} is called with a list of all secret variables (if any) in the
condition of that control structure. When an assignment is encountered, it
uses \ttt{publicNonLocals} to see if there are any secret variables associated
to an enclosing control structure, outside of which the given name has been declared.

At the end of the algorithm, only leaves corresponding to public names are
retained. This is due to the fact that the only time a leak occurs, from the
perspective of this static analysis tool, is if a secret variable eventually
gets indirectly copied to a public variable.

\section{Related Work}
\label{sec:RelatedWork}

The way that the type-level tagging mechanism works is similar to several other
prior works. One such work is McBride's \textit{Kleisli Arrows of Outrageous
Fortune}~\cite{KleisliArrows}. In that paper, an invalid state is unable to be
reached as the result of a similar type-level tag as the \ttt{Sensitivity} tag
in the present paper: the McBride paper uses a type-level tag which indicates
whether a file handle is open or closed. However, significantly more machinery
is involved in the McBride paper, as it is necessary to put such a tag on a
monadic type. While a monadic type is used in the present paper (in the form of
the \ttt{Cmd} type), this type does not use a type-level tag. The creation of
such machinery is the main purpose of the McBride paper.

A significant recent work also related to the use of domain specific languages
with strong, static type systems to control information flow for security
purposes is \textit{Liquid Information Flow Control} by Polikarpova et
al.~\cite{Lifty} In this work, a novel domain specific programming language is
developed which has a static type system that is geared specifically towards
expressing security policies at a type level. As this is done within a static
type system, these invariants are machine-verified at compile-time.  Not only
can these policy violations be detected, the system they develop is also able to
fix leaks.

\section{Future Work}
\label{sec:FutureWork}
A further investigation into the tradeoffs and possible additional features of
the information flow analysis is likely to be useful. This could provide a more
accurate and succinct picture to programmers of where leaks are occurring, or
potentially occurring. It is possible that some of these other leak conditions
could be incorporated into the type system, by exploiting the dependent type
features enabled by singletons~\cite{SingletonsPaper} and, in the future, by
Dependent Haskell.~\cite{DepHaskSpec} Dependent types provide a powerful
technique for writing programs at a type-level, which will be executed at
compile-time by the type checker.~\cite{CertProg, DepHaskSpec} A more
lightweight technique, which might be able to be applied to allow for more
expressive policies to control of information flow within the type system, would
be the ``ghosts of departed proofs'' design technique.~\cite{DepartedProofs}

Another possible, related, future development would be the to attempt to add a
mechanism which automatically fixes some (or all) leaks. Such a system has precendent
in the previously mentioned paper \textit{Liquid Information Flow Control}.~\cite{Lifty}
This could give programmers an additional tool towards ensuring correctness, with
respect to SpectreGuard annotations.

@@

\section{Conclusion}
\label{sec:Conclusion}
